Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-01 paper/disc20clustering.tex
wip
PPL question
Table updated

\title{Improved Bounds for Online Balanced Repartitioning}

@@

\copyrightyear{2020} 
\acmYear{2020} 
\setcopyright{acmlicensed}
\acmConference{PODC '20}{July 29-August 2, 2020}{Toronto, Canada}

@@

\newcommand{\comm}{\textsc{comm}}
\newcommand{\OFF}{\textsc{Off}\xspace}

@@

\maciek{Title discussion: "Online Balanced Graph Partitioning: Understanding the role of clusters". Reason: all our results most notably pertain to the number of clusters.}

@@

An algorithm serves a communication request between two nodes
either \emph{locally} at cost~0
if they are collocated,
or \emph{remotely} at cost~1
if they are located in different clusters.
We refer to these two types of requests as \emph{internal}
and \emph{external} requests, respectively.
\mahmoud{"Before serving a request": does this mean that only after a request arrives and before serving it we migrate? In our algorithms we migrate after serving $\lambda$ requests. Do we have discrepancy with the model?}
Before serving a request,
an online algorithm may perform a \emph{repartition},
%(i.e., \emph{reconfigure}).
i.e.,
it may move (``migrate'') some nodes into clusters different from their current clusters, while respecting the capacity of every cluster. 
Afterwards, 
the algorithm serves the  request.
The cost of migrating a node from one cluster to another
is~$\alpha \in \mathbb{Z}^+$.
For any algorithm $\ALG$,
its cost,
denoted by $\ALG(\sigma)$,
is the total cost of communications and
the cost of migrations performed by $\ALG$ while serving the sequence $\sigma$.

@@

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.5}
    \begin{tabular}{>{\centering\arraybackslash}p{4.5cm}|>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{4.5cm}}
    \rowcolor{gray!50}
    \textbf{Variant} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
      \textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & TBD\hspace{0.3cm}(\S 9.3) \\ 
      \rowcolor{gray!25}
      \textbf{$k=3$}&  $27\cdot \ell $\hspace{0.3cm}(\S 9.3)& $O(\ell)$\hspace{0.3cm}(\S 9.3)\\
      \textbf{$\ell=2$}&  $k+1 = \Omega(k)$\hspace{0.3cm}(\S 9.3)&\hspace{0.3cm}$O(k^2)$\hspace{0.3cm}\cite{repartition-disc}\\
      \rowcolor{gray!25}
      $\ell \geq 3$& $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S 9.3)&$O(k^2 \cdot \ell^2)$\hspace{0.3cm}\cite{repartition-disc} \\
      Perfect partition & $\Omega(k\cdot \ell)$\hspace{0.3cm}(\S 9.3)&$O(k \cdot \ell)$\hspace{0.3cm} (\S 9) \\
    \end{tabular}
                \caption{Overview of the results. Most specific cases are tight. The general case still has a significant gap between the lower bound and the upper bound. If either $k$ or $\ell$ is missing from a variant description, it is assumed to be arbitrary.
                  }
  \label{tab:overview}
  \vspace{-7mm}
\end{table*}

@@

\subsection{Upper bound.}
Now we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
%\maciek{This is not sufficient! We also disallow serving requests remotely.}
where an optimal offline algorithm ($\OPT$) moves to a perfect partition
at the beginning and stays there perpetually.
\maciek{Already mention that the LB is for that scenario}
%We refer this variant as \PPOBRP{}.
The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect partition while not paying too much relative to \OPT.
%
%\maciek{This is not a valid description of the model. We need the assumption that collocation is needed in our model. Otherwise OPT would just serve some requests remotely, and our analysis would not hold.}
The first request between any two non-collocated nodes reveals an edge of the communication graph.
Our online algorithm collocates them immediately and never separates them since by assumption \OPT does the same.
Revealed edges, up to any point in the sequence,
induce some connected subgraphs which grow as components \maciek{rewrite sentence, expand}.
%An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component respecting}.

@@

\section{Improved algorithm for $k=2$}

@@

\maciek{How to distribute algorithms: PPL and DET}

>>>>

\title{Online Balanced Graph Partitioning: Optimizations for Large Clusters}

@@

\copyrightyear{2020} 
\acmYear{2020} 
\setcopyright{acmlicensed}
\acmConference{DISC '20}{October 12-16, 2020}{Freiburg, Germany}

@@

\newcommand{\comm}{\textsc{comm}}
\newcommand{\OFF}{\textsc{Off}\xspace}
\newcommand{\Rep}{\textsc{Rep}}

@@

@@

An algorithm serves a communication request between two nodes
either \emph{locally} at cost~0
if they are collocated,
or \emph{remotely} at cost~1
if they are located in different clusters.
We refer to these two types of requests as \emph{internal}
and \emph{external} requests, respectively.
Before serving a request,
an online algorithm may perform a \emph{repartition},
%(i.e., \emph{reconfigure}).
i.e.,
it may move (``migrate'') some nodes into clusters different from their current clusters, while respecting the capacity of every cluster. 
Afterwards, 
the algorithm serves the  request.
The cost of migrating a node from one cluster to another
is~$\alpha \in \mathbb{Z}^+$.
\maciek{Why restrict $\alpha$? It should just be positive}
For any algorithm $\ALG$,
its cost,
denoted by $\ALG(\sigma)$,
is the total cost of communications and
the cost of migrations performed by $\ALG$ while serving the sequence $\sigma$.

@@

\begin{table*}
  \centering
  \renewcommand{\arraystretch}{1.5}
    \begin{tabular}{>{\centering\arraybackslash}p{4.5cm}|>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}p{4.5cm}}
    \rowcolor{gray!50}
    \textbf{Variant} & \textbf{ Lower bound} &\textbf{Upper bound}\\ \hline 
      \textbf{$k=2$}& 3\hspace{0.3cm}\cite{repartition-disc} & 6\hspace{0.3cm}(\S \ref{sec:k2}) \\ 
      \rowcolor{gray!25}
      \textbf{$k=3$}&  $\ell - 3/2$ \hspace{0.3cm}(\S \ref{sec:lowerbound})& $54\cdot \ell $\hspace{0.3cm}(\S \ref{sec:k3})\\
      $k > 3$ & $(k\cdot \ell - k - \ell) / 2$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$O(k \cdot \ell \cdot \min \{ k \cdot \ell, \Rep(k, \ell) \})$\hspace{0.3cm} (\S \ref{sec:k3}) \\
      \rowcolor{gray!25}
      Perfect partition & $(k\cdot \ell - k - \ell) / 2$\hspace{0.3cm}(\S  \ref{sec:lowerbound})&$2(k -1 )\cdot \ell$\hspace{0.3cm} (\S \ref{sec:ppl}) \\
    \end{tabular}
                \caption{Overview of known results and the results provided in this paper. By $\Rep$ we denote the cost of a single component repartition, and we elaborate on this in Section~\ref{sec:k3}
                  }
  \label{tab:overview}
  \vspace{-7mm}
\end{table*}

@@

\subsection{Upper bound.}
\label{sec:ppl}
Now we give an algorithm for a restricted variant of  \OBRP{}, named \PPOBRP{},
%\maciek{This is not sufficient! We also disallow serving requests remotely.}
where an optimal offline algorithm ($\OPT$) moves to a perfect partition
at the beginning and stays there perpetually.
\maciek{Already mention that the LB is for that scenario}
%We refer this variant as \PPOBRP{}.
The task of an online algorithm for \PPOBRP{} is to recover (or learn) the perfect partition while not paying too much relative to \OPT.
%
%\maciek{This is not a valid description of the model. We need the assumption that collocation is needed in our model. Otherwise OPT would just serve some requests remotely, and our analysis would not hold.}
The first request between any two non-collocated nodes reveals an edge of the communication graph.
Our online algorithm collocates them immediately and never separates them since by assumption \OPT does the same.
Revealed edges, up to any point in the sequence,
induce some connected subgraphs which grow as components \maciek{rewrite sentence, expand}.
%An algorithm that always  keeps nodes belonging to the same component collocated is \emph{component respecting}.

@@

\section{Improved algorithm for $k=2$}
\label{sec:k2}

@@

\subsection{Distributed implementation of $k=2$ algorithm}

Trivial? Counters are local. We migrate between the clusters that have just communicated.

\subsection{Distributed implementation of PPL}

\maciek{How to distribute PPL better?}
